The Hydrogen Epoch of Reionization Array Dish I: Beam Pattern Measurements and Science Implications

### 2015-12-23 DishBeamMeasurements/paper/DishBeamMeasurementsPaper.tex
update to text
smaller delay range on plots, and progress in science implications section

\begin{figure*}
\centering
\includegraphics[width=7in]{dave197_rich195_airy_beams.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspec}
\end{figure*}

@@

\begin{figure*}[h]
\centering
\includegraphics[width=6.7in]{null_expt_rel_beam_maps.pdf}
\caption{null experiment beam ratio plots}
\label{fig:nullexptplots}
\end{figure*}

@@

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.7in]{measured_beams_and_models_maps.pdf}
\caption{null experiment beam ratio plots}
\label{fig:measuredbeammaps}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=6.7in]{measured_beams_and_models_slices.pdf}
\caption{null experiment beam ratio plots}
\label{fig:measuredbeamslices}
\end{figure*}

@@

\subsection{Collecting Area}

@@

 \begin{table}[h]
 \caption{ \label{table:collectingareatable}Collecting area (m$^2$) of measured 137\,MHz beams and corresponding power spectrum SNR for HERA-127.}
\begin{tabular}{| l | l | l |}
\hline
Beam & $A_\text{eff}$ (m$^2$) & SNR (pess, mod, opt)\\
\hline
  Airy pattern & 155 & 9.1, 11.0, 37.2  \\
  %CST sim & 67.0 &  \\
  %Measured, feed at 4\,m & 44.3 &  \\
    Measured, feed at 5.3\,m & 97.9 & 8.2, 8.3, 29.2 \\
    Measured, feed at 5\,m & 82.6 & 6.8, 6.9, 26.5 \\
    Measured, feed at 4.5\,m & 73.6 & 6.4, 6.5, 24.8 \\ 
  \hline
\end{tabular}
\end{table}

@@

\section{Foreground Leakage and Residuals}
\label{sec:foregroundleakage}

\begin{figure*}[h]
\includegraphics[width=6.7in]{delay_spectra_dave_rich_airy_beams_manyLSTs.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspecmanyLSTs}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{gsm_kelvin_LST_2deg_and_62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:gsmplots}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{delay_spectra_dave_rich_airy_beams_LST2deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspecbadLST}
\end{figure*}

\begin{figure*}[h]
\includegraphics[width=6.7in]{delay_spectra_dave_rich_airy_beams_LST62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspecgoodLST}
\end{figure*}

@@

add stuff here

>>>>

\begin{figure*}
\centering
\includegraphics[width=7in]{dave197_rich195_airy_beams.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:modelbeams}
\end{figure*}

@@

\begin{figure*}[h]
\centering
\includegraphics[width=6.5in]{null_expt_rel_beam_maps.pdf}
\caption{null experiment beam ratio plots}
\label{fig:nullexptplots}
\end{figure*}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=6.5in]{measured_beams_and_models_maps.pdf}
\caption{null experiment beam ratio plots}
\label{fig:measuredbeammaps}
\end{figure*}

\begin{figure*}[t]
\centering
\includegraphics[width=6.5in]{measured_beams_and_models_slices.pdf}
\caption{null experiment beam ratio plots}
\label{fig:measuredbeamslices}
\end{figure*}

@@

@@

\subsection{Sensitivity}

@@

 \begin{table}[h]
 \caption{ \label{table:collectingareatable}Collecting area (m$^2$) of measured 137\,MHz beams and corresponding power spectrum SNR for HERA-127.}
\begin{tabular}{| l | l | l |}
\hline
Beam & $A_\text{eff}$ (m$^2$) & SNR (pess, mod, opt)\\
\hline
  Airy pattern & 155 & 9.1, 11.0, 37.2  \\
  %CST sim & 67.0 &  \\
  %Measured, feed at 4\,m & 44.3 &  \\
    Measured, feed at 5.3\,m & 97.9 & 8.2, 8.3, 29.2 \\
    Measured, feed at 5\,m & 82.6 & 6.8, 6.9, 26.5 \\
    Measured, feed at 4.5\,m & 73.6 & 6.4, 6.5, 24.8 \\ 
  \hline
\end{tabular}
\end{table}\

NEED TO TALK ABOUT 21CMSENSE AND THE SNR CALCULATIONS

@@

\section{Implications for Foreground Suppression}
\label{sec:foregrounds}

ALSO MENTION DEPENDENCE ON BASELINE DIRECTION
DISCUSS DELAY CLEANING TO THE 2min THERMAL NOISE LEVEL, THEN DELAY TRANSFORM AGAIN WITH REDUCED RESOLUTION, MEANING A LARGER BH WINDOW, AND THUS MORE LEAKAGE, THESE EFFECTS ARE DISCUSSED IN NITHYA'S PAPER, here we are focused on intrinsic delay distribution of foregrounds

We simulate the power spectra of foregrounds with different primary beam models under various observing conditions. Broadly, beam models with more response near the horizon see smooth spectrum sources with more frequency dependence (higher delay), appearing closer to the nominally clean EOR window, making it more at risk of leaking power inside. We use the per-baseline approach of XX and XX, simulating visibilities measured by specific baselines as a function of frequency, then computing the Fourier transform over frequency (the delay transform) and normalizing the result into a cosmological power spectrum \citep{perbaselinetechnique}.

We simulate visibilities for three beam models (HFSS, CST, and an Airy pattern for a 14\,m aperture) at various LSTs, modeling the sky as the sum of the Global Sky Model \citep{gsm} and a point source catalog (XX). We use a frequency spacing of 1\,MHz, sufficient to characterize delays within and just outside of the horizon limits on both baseline lengths we are concerned with, 14\,m and 42\,m. We use a total bandwidth of 80\,MHz (40\,MHz after applying the Blackman-Harris window) centered on 150\,MHz. This bandwidth is larger than the 10\,MHz thought to be safe from signal evolution over redshift, but is the bandwidth used in the power spectrum analysis of XX,XX to perform a high delay resolution empirical foreground subtraction (delay-space CLEANing) as described by XXX. We normalize the delay spectra into power spectra following   \citet{nithya15}.

Nithya et al (submitted) detail the implications of these foreground delay spectra for suppression strategies. We are concerned here with the role of the beam in these predictions, in particular what our beam model and its uncertainties imply about foreground suppression strategies. We begin by simulating delay spectra, plotted in Figure \ref{fig:delayspec} (top panel) at various LSTs as outlined above using the nominal HFSS beam. As all these LSTs are high galactic latitudes far from the galactic center, the total visibility power (the level of the zero delay mode) varies by only an order of magnitude over these LSTs both on the 14\,m (left panel) and 42\,m (right panel). However the horizon limit at negative delay (corresponding to the western horizon) shows a peak that varies by three orders of magnitude on the 14\,m baseline and two orders of magnitude on the 42\,m baseline, showing the stark difference in pitchfork brightening when the galaxy is just above versus just below the horizon.

In this figure we perform the approximate conversion from delay to k//, which we plot as a second x-axis at the top of the plot. For these short baselines, k// may be converted to k by adding kperp=0.005 or 0.02 in quadrature for these baselines at this redshift. These numbers are small compared to the range of k// plotted, and thus we interpret the k// axis as simply the k axis, and plot a 1D model power spectrum computed using 21cmFast as a dotted line for comparison.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_manyLSTs.pdf}
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_LST2deg.pdf}
\includegraphics[width=6in]{delay_spectra_dave_rich_airy_beams_LST62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:delayspec}
\end{figure*}

We study the differences between our three beam models on two of these LSTs, one with maximal horizon brightening ($2^\circ$), and one with minimal horizon brightening ($62^\circ$). Figure \ref{fig:gsmplots} shows the sine-projected Global Sky Model in horizontal coordinates with units of Kelvin for these two LSTs. Dashed lines mark zenith angles 20$^\circ$, 40$^\circ$, 60$^\circ$, 80$^\circ$. These plots confirm that the large negative delay peak at both baseline's horizon limits at the 2 degree LST originates from the center of the galaxy being just above the horizon. In contrast several hours later, the galactic center is fully below the western horizon, leaving only a slight brightening near the eastern horizon due to the much weaker galactic anticenter.

\begin{figure*}[h]
\centering
\includegraphics[width=6in]{gsm_kelvin_LST_2deg_and_62deg.pdf}
\caption{Simulated delay spectra for the GSM and point sources at $0^\circ$ (top pair) and $60^\circ$ (bottom pair) LST for EW baselines of length 14\,m (left pair) and 42\,m (right pair). The beams are are assumed constant over frequency.}
\label{fig:gsmplots}
\end{figure*}

We turn now to the differences in foreground delay spectra in these two observing conditions with different beam models. Figure \ref{fig:delayspec} (middle panel) shows the delay spectra for all three beam models at $2^\circ$ LST, which has the worst horizon brightening. We use both numerical beam models with 5\,m feed heights, as well as the Airy pattern for a 14\,m aperture. The delay spectra for these beams agree out to delays of 10--20\,ns on the 14\,m baseline and 30--40\,ns on the 62m baseline. These limits suggest that the beams track each other fairly well out to 15--20$^\circ$ from zenith, beyond which they cease to be correlated. This is essentially what is observed in Figure \ref{fig:measuredbeamslices} (middle panel). At larger delays, especially near the negative delay horizon limit, all three model delay spectra diverge due to the significant edge brightening which effectively discriminates between these models. The CST, HFSS, and Airy beams reach roughly -35\,dB, -38\,dB, and -50\,dB at the horizon (Figure \ref{fig:modelbeams}), consistent with the fact that the CST beam has the largest horizon brightening, followed by the HFSS beam, and then the Airy beam. This is seen in the delay spectra from both baseline lengths, though the edge brightening is much more significant in the longer baseline where it is less diluted by the zero delay emission.

In contrast, all three model agree much more closely when there is little or no edge brightening as in Figure \ref{fig:measuredbeamslices} (middle panel) where we plot the delay spectra for all three beams for 62$^\circ$ LST. There is still a modest edge brightening observed here, a flattening off near the horizon on the short baseline and a small peak on the long baseline, both due to the increasing solid angle per zenith angle at larger zenith angles. However as the emission at this LST at large delays is roughly the same temperature as emission from everywhere else on the sky, washing out the difference between the three beam models.

@@

%- power spectrum analyses by first generation instruments are ongoing, but face many challenges
%- that’s why HERA was designed to be optimized for 21cm

%- in this paper we have characterized the HERA dish’s angular response
%  down to 60deg zenith angle, -35dB from zenith
%- confirms models to exquisite fidelity unachievable with celestial sources
%- the level of model disagreement also gives us a handle on antenna-to-to antenna variation

%- given the latest understanding of instrument chromaticity and
%  foreground isolation (ie, Nithya), we simulate the effects of these
%  beams on science
 %   - taking into account uncertainties in near-horizon beam response unprobed by our measurements
%- we make predictions for the level of horizon brightening at different LSTs, and discuss the uncertainties due to uncertain beam models (horizon brighting is an exquisite of the low elevation beam response, though an integrated statistic)
% horizon brightening will dominated leakage into the outer window modes

% frequency dependence of the beam pattern and overall gain (due to reflections) likely dominates leakage deep within the window, but the other papers in this series demonstrate that this leakage level is within our spec

   % - all this highlight the need for detailed antenna characterization
     % and simulation of the implications for science analyses  in 21cm observatories
